\propertitle

### 2019-12-15 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-15T18:01:21+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Inferences about large neuronal populations}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\begin{figure}[!p]
\centering
\subcaption[]{\label{fig:plots_cellsubsets_a}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper3_gridvsnogrid.pdf}%
\\[2em]%
\subcaption[]{\label{fig:plots_cellsubsets_b}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper3_gridconv.pdf}%
\\[2em]%
\caption{\subcaptionref{fig:plots_cellsubsets_a} Maximum-entropy
  distributions for larger subpopulations or grid and non-grid cells.
  \subcaptionref{fig:plots_cellsubsets_b} Convolution of the two
  subpopulation distributions; it would be equal to the total distribution
  if the two distributions were independent.}
\label{fig:plots_cellsubsets}
\end{figure}
% maxent_solutions_different_moments_partial.nb
Here is a toy example of possible uses of this model, based on the first
data set. We could be interested in the hypothesis that two distinct cell
types or assemblies be present in the region where the recording was made.
Finding a larger-population distribution with two peaks, as in
\fig~\ref{fig:plots_5mom_a}, would provide some evidence for this
hypothesis. We might have reasons for suspecting that a specific set of the
sampled neurons is of the first type, and the remaining of the second type.
In the case of the first data set, 27 of the 65 sampled neurons were
identified as grid cells belonging to 3--4 functional modules
\citep{stensolaetal2012}. Could the two peaks in the distribution of
\fig~\ref{fig:plots_5mom} reflect the activities of grid versus non-grid
cells? We apply the model to these two sets of neurons individually, using
$\yNg=4\,150$ for the grid set and $\yNng=5\,850$ for the non-grid set, to
reflect their proportions (27/65 and 38/65) in the recorded sample. The
results are shown in \fig~\ref{fig:plots_cellsubsets_a}. The distribution
for the larger population of grid cells (\textcolor{mygreen}{green
  triangles}) seems to have one broad peak, close to the first peak of the
larger-population distribution. The distribution for the larger population
of non-grid cells (\textcolor{myred}{red circles}) has two peaks instead,
roughly at the same population-averaged activities as the peaks of the
larger-population distribution (\textcolor{myblue}{blue curve}) but closer
in height. It would thus seem that the population of grid cells is
contributing to the first mode of the larger population, but it is not its
sole contributor. We can also assess if the distributions for the two sets
of neurons are independent. If they were independent, the larger-population
distribution would be given by their convolution:
\begin{equation}
  \label{eq:convolution}
  P_{\text{full}}(\yA) = \sum_{\yA'} P_{\text{grid}}(\yA')\;
  P_{\text{non-grid}}(\yA-\yA') %, \qquad \yA \in \set{1,\dotsc,10\,000}
\end{equation}
where the index $\yA'$ runs from $\max(0,\yA-\yNng)$ to $\min(\yA,\yNg)$.
But this is not the case: 
figure~\ref{fig:plots_cellsubsets_b} shows that such convolution
(\textcolor{black}{black diamonds}) is quite different from the
larger-population distribution (\textcolor{myblue}{blue curve}): the two
peaks of the former are closer in position and height than those of the
latter. The population distributions of grid and non-grid cells are
therefore \emph{not} independent: knowledge of the activity of either set
gives us some information about the activity of the other.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-15T18:25:32+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Inferences about large neuronal populations}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

% maxent_solutions_different_moments_partial.nb
Here is a toy example of possible uses of this model, based on the first
data set. We could be interested in the hypothesis that two distinct cell
types or assemblies be present in the region where the recording was made.
Finding a larger-population distribution with two peaks, as in
\fig~\ref{fig:plots_5mom_a}, would provide some evidence for this
hypothesis. We might have reasons for suspecting that a specific set of the
sampled neurons is of the first type, and the remaining of the second type.
In the case of the first data set, 27 of the 65 sampled neurons were
identified as grid cells belonging to 3--4 functional modules
\citep{stensolaetal2012}. Could the two peaks in the distribution of
\fig~\ref{fig:plots_5mom} reflect the activities of grid versus non-grid
cells? We apply the model to these two sets of neurons individually, using
$\yNg=4\,150$ for the grid set and $\yNng=5\,850$ for the non-grid set, to
reflect their proportions (27/65 and 38/65) in the recorded sample. The
results are shown in \fig~\ref{fig:plots_cellsubsets_a}.
\begin{figure}[!p]
\centering
\subcaption[]{\label{fig:plots_cellsubsets_a}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper3_gridvsnogrid.pdf}%
\\[2em]%
\subcaption[]{\label{fig:plots_cellsubsets_b}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper3_gridconv.pdf}%
\\[2em]%
\caption{\subcaptionref{fig:plots_cellsubsets_a} Maximum-entropy
  distributions for larger subpopulations or grid and non-grid cells.
  \subcaptionref{fig:plots_cellsubsets_b} Convolution of the two
  subpopulation distributions; it would be equal to the total distribution
  if the two distributions were independent.}
\label{fig:plots_cellsubsets}
\end{figure}
The distribution for the larger population of grid cells
(\textcolor{mygreen}{green triangles}) seems to have one broad peak, close
to the first peak of the larger-population distribution. The distribution
for the larger population of non-grid cells (\textcolor{myred}{red
  circles}) has two peaks instead, roughly at the same population-averaged
activities as the peaks of the larger-population distribution
(\textcolor{myblue}{blue curve}) but closer in height. It would thus seem
that the population of grid cells is contributing to the first mode of the
larger population, but it is not its sole contributor. We can also assess
if the distributions for the two sets of neurons are independent. If they
were independent, the larger-population distribution would be given by
their convolution:
\begin{equation}
  \label{eq:convolution}
  P_{\text{full}}(\yA) = \sum_{\yA'} P_{\text{grid}}(\yA')\;
  P_{\text{non-grid}}(\yA-\yA') %, \qquad \yA \in \set{1,\dotsc,10\,000}
\end{equation}
where the index $\yA'$ runs from $\max(0,\yA-\yNng)$ to $\min(\yA,\yNg)$.
But this is not the case: 
figure~\ref{fig:plots_cellsubsets_b} shows that such convolution
(\textcolor{black}{black diamonds}) is quite different from the
larger-population distribution (\textcolor{myblue}{blue curve}): the two
peaks of the former are closer in position and height than those of the
latter. The population distributions of grid and non-grid cells are
therefore \emph{not} independent: knowledge of the activity of either set
gives us some information about the activity of the other.
